---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
author: "Hankang"
date: "December 11, 2014"
---
```{r setup}
# set global chunk options: 
library(knitr)
library(plyr)
library(dplyr)
library(ggplot2)
opts_chunk$set(fig.align='center', echo = TRUE)
```

## 1. Loading and preprocessing the data

```{r Unzip the "activity.zip" file and load it to df}
unzip("activity.zip")
df <-  read.csv("activity.csv", sep = ",")
```

## 2. What is mean total number of steps taken per day?

* 1 Make a histogram of the total number of steps taken each day
```{r, Load library "dplyr" then use ddply to get the total steps}
df.steps <- ddply(df, .(date), summarize, total = sum(steps, na.rm = T))
```

```{r, histogram of the total number of steps taken each day, fig.height = 4}
df.steps.hist <- as.POSIXlt(with(df.steps,rep(date,total)))
hist(df.steps.hist, "weeks", breaks = 61, freq = T, xlab = "Date",main = "Histogram of total steps in Oct and Nov", xaxt = "n")
axis.POSIXct(1, at=seq(as.POSIXlt("2012-10-01"), as.POSIXlt("2012-11-30"), by="1 weeks"), format="%Y-%m-%d",las=2, cex.axis = 0.5)
```

* 2 Calculate and report the mean and median total number of steps taken per day

```{r, Calculate the mean and media}
df.summary <- ddply(df, .(date), summarize, mean = mean(steps, na.rm = T), median = median(steps, na.rm = T))
print(df.summary)
```
wired summary... possibly too manmy zeros in the data, try no include "0" or "NA" ?

```{r, Remove "0" and "NA", then calcluate the mean and media}
df.NA = df[df$steps != 0 & !is.na(df$steps),]

df.summary.NA <- ddply(df.NA, .(date), summarize, mean = mean(steps, na.rm = T), median = median(steps, na.rm = T))
print(df.summary.NA)
```

**NOTE: **_Much better, but we can see we only have 53 dates data!_

## What is the average daily activity pattern?

* 1 Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, time series plot}
df.average <- ddply(df, .(interval), summarize, mean = mean(steps, na.rm = T))

with(df.average, plot(interval, mean, type = "l",xlab = "Interval (mins)", ylab = "Averaged number of steps taken",main = "Averaged daily activity pattern"))
```

* 2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, maximum number of steps}
with(df.average, interval[which(mean == max(mean))])
```
## Imputing missing values

* 1 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)

```{r, total number of missing values}
NAtotal <- sum(is.na(df$steps))
print(NAtotal)
```
* 2 Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

    **Using mean for that 5-minute interval**

* 3 Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, attach "dplyr" package and use chaining}

temp <- merge(df, df.average, by = "interval")
temp$steps[is.na(temp$steps)] <- temp$mean[is.na(temp$steps)]

imputed <- temp %>%
    arrange(date, interval) %>%
    select(steps, date, interval)

imputed.steps.total <- ddply(imputed, .(date), summarize, 
                       total = sum(steps))
```

* 4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r histogram of imputed table}
imputed.steps.hist <- as.POSIXct(rep(imputed.steps.total$date,
                      imputed.steps.total$total))

hist(imputed.steps.hist, "weeks", breaks = 61, freq = T, xlab = "Date",
     main = "Histogram of total steps in Oct and Nov (imputed)", xaxt = "n")
axis.POSIXct(1, at=seq(as.POSIXlt("2012-10-01"), as.POSIXlt("2012-11-30"), 
                       by="1 weeks"), format="%Y-%m-%d",las=2, cex.axis = 0.5)
```

_Mean and media_

```{r}
## report mean and media
imputed.summary <- ddply(imputed , .(date), summarize, 
                 mean = mean(steps, na.rm = T), 
                 median = median(steps, na.rm = T))
print(imputed.summary)
# remove zero
imputed1 = imputed[imputed$steps != 0,]

imputed.summary1 <- ddply(imputed1, .(date), summarize, 
                  mean = mean(steps, na.rm = T), 
                  median = median(steps, na.rm = T))
print(imputed.summary1)
```

_visulize the difference_
```{r, visulize the difference}
h1 <- hist(df.steps.hist, breaks = 61, freq = F, plot = F)
h2 <- hist(imputed.steps.hist, breaks = 61, freq = F,xaxt = "n",
           plot = F)
plot(h1, col=rgb(0,0,1,1/8), xaxt = "n",main="Comparison", xlab="Date", ylab="Frequency")
plot(h2, col=rgb(1,0,0,1/8), add=T)

axis.POSIXct(1, at=seq(as.POSIXlt("2012-10-01"), as.POSIXlt("2012-11-30"), 
                      by="1 weeks"), format="%Y-%m-%d",las=2, cex.axis = 0.5)
legend("topleft", c("original data","imputed", "overlap"), 
       fill = c(rgb(0,0,1,1/8), rgb(1,0,0,1/8),rgb(1,0,1,1/8)))

```
**Explanation:**
The `NA` was replaced by the mean of 5-minute interval which rised the interval value of those contained `NA` but remained those did not contain `NA` the same. That's why most of the histogram is overlaped.

## Are there differences in activity patterns between weekdays and weekends?

* 1 Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r weekday data frame}
weekdays <- as.POSIXlt(df$date)$wday
df.weekday <- cbind(df, weekdays)

df.weekday$weekdays[which(df.weekday$weekdays %in% c(1:5))] <- "weekday"
df.weekday$weekdays[which(df.weekday$weekdays %in% c(0,6))] <- "weekend"
df.weekday$weekdays <- as.factor(df.weekday$weekdays)

```

* 2 Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r, panel plot}
weekday.summary <- ddply(df.weekday, .(interval,weekdays), summarize, 
                         mean = mean(steps, na.rm = T))
qplot(interval, mean, data = weekday.summary, facets = weekdays ~ ., geom = "path",
      ylab = "Averaged number of steps taken")
```